---
title: "Final"
author: "mwebers2"
date: "12/6/2019"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(shiny)
library(readr)
library(dplyr)
ui <- fluidPage(
  titlePanel("OkCupid"),
      tabsetPanel(
        
        tabPanel("Body Type",
           fluidRow(sidebarPanel(
                    h4("Body Type"),
                    selectInput("Characteristic", "characteristic",
                                  c("gender", "drinks", "drugs", "diet"), multiple = TRUE),
                    mainPanel(plotOutput("hist", height = 500))))), 
        
        tabPanel("Age",
           fluidRow(sidebarPanel(width = 3,
                    sliderInput("age", "Age:", min = 18, max = 110, value = c(25, 30)),
                    mainPanel(plotOutput("hist", height = 500))))),
        
        
        tabPanel("Word Map", 
                 fluidRow(sidebarPanel(
                    h4("Body Type"),
                    selectInput("Characteristic", "characteristic",
                                  c("age","gender"), multiple = TRUE),
                    mainPanel(plotOutput("wordMap"))))),
        
        tabPanel("Word Count",
                 fluidRow(
                   column(width = 3,selectInput("menu","Choose a variable:",c("smokes","drinks","body_type"))),
                   column(width = 9, plotOutput("wordCount")))),
        
        tabPanel("Height",plotOutput("heightPlot")),
        tabPanel("No answer",plotOutput("noAnsPlot"))))

server <- function(input, output) {
  
    output$bodyPlot <- renderPlot({
      #plotting goes here
    })
    
    output$agePlot <- renderPlot({
      #go plot here
    })
    
    output$wordMap <- renderPlot({
      #just to hold space
    })
    
    output$wordCount <- renderPlot({
      library(tidyverse)
      dating <- read.csv("profiles.csv")
      attach(dating)
      essays <- dating[,7:16]
      essays <- apply(essays,2,as.character) 
      bruh <- apply(essays,2,nchar)
      sums <- apply(bruh,1,sum)
      dating$nchar <- sums
      select <- input$menu
      if (select == "smokes") {
        ok <- ggplot(data = dating) +
          geom_histogram(mapping = aes(x = nchar, fill = smokes)) +
          xlim(0,20000)
      }
      if (select == "drinks") {
        ok <- ggplot(data = dating) +
          geom_histogram(mapping = aes(x = nchar, fill = drinks)) +
          xlim(0,20000)
      }
      if (select == "body_type") {
        ok <- ggplot(data = dating) +
          geom_histogram(mapping = aes(x = nchar, fill = body_type)) +
          xlim(0,20000)
      }
      ok
    })
    
    output$heightPlot <- renderPlot({
      
    })
    
    output$noAnsPlot <- renderPlot({
      
    })
}
# Run the application 
shinyApp(ui = ui, server = server, options = list(height = 500))
```

View(dating)

essays <- dating[,7:16]
dating[,7:16] <- lapply(dating[,7:16],as.character)
dating$word_count <- apply(dating[,7:26],2,nchar)
men <- filter(dating, sex == "m")
men_total <- mean(men$word_count, na.rm = T)
women <- filter(dating, sex == "f")
women_total <- mean(women$word_count, na.rm = T)


summary(dating$age)
factor(dating$drugs)
table(dating$drugs)
ggplot(data = dating) +
  geom_bar(mapping = aes(x = drugs)) +
  coord_flip()
ggplot(data = dating) + 
  geom_bar(mapping = aes(x = body_type)) +
  coord_flip()
ggplot(data = dating) +
  geom_bar(mapping = aes(x = offspring)) + 
  coord_flip()
#body type by age, gender, diet, smokes, drugs, height, ethnicity
#orientation by religion, sex, star sign
#sex by age

white <- filter(dating, ethnicity == "white")
jacked <- filter(white, body_type == "jacked")
View(jacked$essay0)

length(dating$age)

goal: look at OkCupid Cupid to find partner
ask Chris
are most people on OkCupid educated? find yes or no
filters by education
if you want "x", you should look for "y"

